---
title: "Covid91 vaccine study FinalFall2022"
author: "America Olivo"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="finalFall2022.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
M = filter(FinalData, sex =="M")
F = filter(FinalData, sex =="F")
gay = filter(FinalData, LGBTQ =="gay")
Druguser = filter(FinalData, DrugUser =="yes")
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

## Males

```{r}
basic <- M %>% group_by(infected,treatment) %>% summarize(count=n()) %>%  mutate(prcnt=count/sum(count))
basicC <- ggplot(basic,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")
basicCC <- ggplot(basic,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```


This bar graph shows the population of men who got covid by either a placebo or vaccine. The placebo had more sick men than the actual vaccine. Over 60% of men contracted covid after the placebo and the low of about 30% of men who got sick after the vaccine. Men were twice as likely to get sick/covid from the placebo than the vaccine.In the second graph,the x-axis represents  if a person got covid-19 or not. The y-axis represents the percentage of people that got covid-19 after taking the the placebo or taking the vaccine.


### Numerical results


```{r}
table1 <- xtabs(~infected + treatment, data=M)
rowPerc(table1)
colPerc(table1)
```


The table shows the percentages of men that got covid-19 after taking the placebo and taking the vaccine. 68% of men who recieved the placebo got sick while the vaccine had a 32% rate of men that got sick after. 


### Inferential 


```{r}
chisq.test(table1)
chisqtestGC(table1)
Mfish <- fisher.test(table1)
Mfish
```


The pvalue is shown to be 2.364e-12(0.00000000002364) is a small number which means these results are liable and not likely to occur by chance. 

...



```{r}
effect <- Mfish$p.value/(Mfish$p.value +1)
effectivenessM = Mfish$estimate/(Mfish$estimate +1)*100
print(effectivenessM)
```


## Effectiveness Confidence Interval `r low`  to `r hi`





## Females


## LGBTQ



## Druggies


# Overall Results and Conclusions