---
title: "Covid91 vaccine study FinalFall2022"
author: "America Olivo"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="finalFall2022.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
M = filter(FinalData, sex =="M")
F = filter(FinalData, sex =="F")
gay = filter(FinalData, LGBTQ =="gay")
Druguser = filter(FinalData, DrugUser =="yes")
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

## Males

```{r}
basic <- M %>% group_by(infected,treatment) %>% summarize(count=n()) %>%  mutate(prcnt=count/sum(count))
basicC <- ggplot(basic,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")
basicCC <- ggplot(basic,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```


This bar graph shows the population of men who got covid by either a placebo or vaccine. The placebo had more sick men than the actual vaccine. Over 60% of men contracted covid after the placebo and the low of about 30% of men who got sick after the vaccine. Men were twice as likely to get sick/covid from the placebo than the vaccine.In the graph,the x-axis represents  if a person got covid or not. The y-axis represents the percentage of people that got covid after taking the the placebo or taking the vaccine.


### Numerical results


```{r}
table1 <- xtabs(~infected + treatment, data=M)
rowPerc(table1)
colPerc(table1)
```


The table shows the percentages of men that got covid after taking the placebo and taking the vaccine. 68% of men who received the placebo got sick while the vaccine had a 32% rate of men that got sick after. 


### Inferential 


```{r}
chisq.test(table1)
chisqtestGC(table1)
Mfish <- fisher.test(table1)
Mfish
```


The pvalue is shown to be 2.364e-12(0.00000000002364) is a small number which means these results are liable and not likely to occur by chance. We see that sex does play a role in the deaths of individuals.



```{r}
effect <- Mfish$p.value/(Mfish$p.value +1)
effectivenessM = Mfish$estimate/(Mfish$estimate +1)*100
print(effectivenessM)
```

The odds ratio is how effective the vaccine is. The odds ratio seen above is 69.


```{r}
low=(Mfish$conf.int[1]/(Mfish$conf.int[1] +1))*100
hi=(Mfish$conf.int[2]/(Mfish$conf.int[2] +1))*100
options(digits=2)
```




## Effectiveness Confidence Interval 

`r low`  to `r hi`

Up is the effectiveness for the covid vaccine for males. If the effectiveness is above 50 that means that the vaccine is effective. From the interval we can see that the vaccine was effective among men. The effectiveness confidence interval for men is 63.401 to 73.347. This proves the vaccine was effective among men. 



## Females

### Barplot

```{r}
basic<- F %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(basic,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(basic,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

This barplot represents the population of females and the relationship of getting covid after by either taking the vaccine or the placebo. It shows that more women who were only given the vaccine contracted covid than those who took the placebo. 58% of women got covid with the vaccine while only approximately 42% got covid from the placebo. In the bar plot, the x-axis represents whether a person got covid-19 or not. The y-axis represents the percentage of people that got covid after taking the  placebo or  the vaccine. 


### Numerical results

```{r}
table1 <- xtabs(~infected + treatment, data=F)
rowPerc(table1)
colPerc(table1)
```

These numerical results show the  percentages of women that got covid-19 after having the placebo and the vaccine.The women who took the vaccine and got covid was 57% and 43% of those took the placebo. This proves that the placebo is more effective than the vaccine.The women that took the vaccine were more likely to get covid than those who took the placebo. The odds of getting covid with the vaccine was almost twice as much of women getting sick than with the placebo. 


### Inferential 
```{r}
chisq.test(table1)
chisqtestGC(table1)
Ffish <- fisher.test(table1)
Ffish
```
The p-value displayed above is small 0.005 




```{r}
effect <- Ffish$p.value/(Ffish$p.value +1)
effectivenessF = Ffish$estimate/(Ffish$estimate +1)*100
print(effectivenessF)
```
The odds ratio is how effective the vaccine is 43.

```{r}
low=(Ffish$conf.int[1]/(Ffish$conf.int[1] +1))*100
hi=(Ffish$conf.int[2]/(Ffish$conf.int[2] +1))*100
options(digits=2)
```

## Effectiveness Confidence Interval 

`r low`  to `r hi`


The effectiveness for covid for females. If the effectiveness is above 50% that means that the vaccine is effective. Based on the interval, we see that the vaccine is not effective in protecting women from covid.The effectiveness confidence interval for women is 37.34 to 47.86.


## LGBTQ

### Barplot

```{r}
basic <- gay %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))
basicC <- ggplot(basic,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")
basicCC <- ggplot(basic,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```


This barplot shows the population of gay people and getting covid by taking either the vaccine or the placebo. As shown in the barplot, it shows that approximately 15% of gay people got covid after taking the placebo and 85% of gays got covid after taking the vaccine. 


### Numerical results

```{r}
table1 <- xtabs(~infected + treatment, data=gay)
rowPerc(table1)
colPerc(table1)
```


This table above shows the infected gay people who took the placebo was a low 13% and those who took the vaccine was at a high of 87%. 



### Inferential 


```{r}
chisq.test(table1)
chisqtestGC(table1)
gayfish <- fisher.test(table1)
gayfish
```

 p-value is 0 which is low. 
 
 
```{r}
effect <- gayfish$p.value/(gayfish$p.value +1)
effectivenessgay = gayfish$estimate/(gayfish$estimate +1)*100
print(effectivenessgay)
```

The odds ratio is how effective the vaccine is 14. Very low indicating it does not work well. 


```{r}
low=(gayfish$conf.int[1]/(gayfish$conf.int[1] +1))*100
hi=(gayfish$conf.int[2]/(gayfish$conf.int[2] +1))*100
options(digits=2)
```



## Effectiveness Confidence Interval

`r low`  to `r hi`

The effectiveness confidence interval for the people who identified as gay is 6.201 to 25.6132 


## Druggies


### Barplot


```{r}
basic <-  Druguser %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))
basicC <- ggplot(basic,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")
basicCC <- ggplot(basic,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```


This barplot shows the population of people who are  drug users and having covid while taking either the vaccine or the placebo. Around 18% of drug users when taking the placebo, got covuid. The other 82% of the drug users got covid after having the vaccine.The drug users that took the vaccine were way more likely to get covid than the placebo.In the bar plot, the x-axis represents whether a person got covid-19 or not. The y-axis represents the percentage of people that got covid after taking the  placebo or  the vaccine. 



### Numerical results

```{r}
table1 <- xtabs(~infected + treatment, data= Druguser)
rowPerc(table1)
colPerc(table1)
```


The numerical results above show the  percentages of drug users that got covid after taking the placebo and the vaccine. The percentage of drug users that took the placebo and got covid was 17% while 83% of drugusers got covid after taking the vaccine. The drug users that took the vaccine were more likely to get than those taking the placebo.


### Inferential 


```{r}
chisq.test(table1)
chisqtestGC(table1)
drugfish <- fisher.test(table1)
drugfish
```


As shown above the p-value is 0 which is low. The Fishers's Exact Test shows that the odds ratio is 0.21.drug users who got the vaccine were about 4-5 times more likely to get covid than those drug users who tookt he placebo. 


```{r}
effect <- drugfish$p.value/(drugfish$p.value +1)
effectivenessdrug = drugfish$estimate/(drugfish$estimate +1)*100
print(effectivenessdrug)
```

the odds ratio is 17 very low meaning it did not work well. 


```{r}
low=(drugfish$conf.int[1]/(drugfish$conf.int[1] +1))*100
hi=(drugfish$conf.int[2]/(drugfish$conf.int[2] +1))*100
options(digits=2)
```


## Effectiveness Confidence Interval 

`r low`  to `r hi`

the effectiveness for the covid for drugusers is 8.571 to 29.4733. Taking the vaccine increases the contraction rate of covid among drugusers. Those didnt take the vaccine were better off without it. 

# Overall Results and Conclusions

The covid vaccine was effective in protecting men against it. The covid vaccince  effectiveness was `r effectivenessM`







